Linear Ordered Attribute Grammar scheduling using SAT-solving

### 2014-10-18 paper/paper.tex
Add a statement about the number of visits found and how it related to previous known results.
Changed the paragraph about input and output attributes

The figure shows what is called the \emph{production dependency graphs}. For a production two different types of attributes can be distinguished: \emph{input} attributes, which are attributes for which the value is provided, and \emph{output} attributes, which are values that need to be defined as part of the production relating parent and children. The input attributes of a production, coloured gray in the picture, are the inherited attributes of the production itself, because their values are provided by the parent node, and the synthesized attributes of all its children. The output attributes, coloured white in the picture, are the synthesized attributes of the parent and the inherited attributes of the child.

Although we talk about dependency graphs in the context of scheduling, we actually draw the edges in the other direction. The edges in the picture represent data flow, which from an attribute grammar perspective is much more intuitive. For example, in the \emph{Bin} production the attribute a1 of the child \emph{r} is computed from a1 from the parent and a3 from the child \emph{l}. The edges are thus always directed from an input attribute to an output attribute, and the actual dependency graph can be obtained by reversing all edges.

@@

@@

To compute this, we introduce a SAT variable $E(a)$ for every attribute $a$, that expresses whether or not $a$ is in $E$ or not. We set $E(a)$ to be true for the initial eager attribute $a$. We go over the graphs for the non-terminals and production rules, and generates constraints that express that whenever $a$ points to $b$ and we have $E(b)$, then we also need $E(a)$.

@@

We have run this algorithm on ever output attribute of the top-level non-terminal of all our examples. For our largest grammar, the hardest output to compute took 1 second. So, while a harder optimisation than the previous one, it is very doable in practice.

@@

In the translation into the SAT-problem we have used a novel technique for ruling out cycles in the SAT problem using chordal graphs. The technique was inspired by existing work on chordility for expressing equality logic \cite{bryant-chordal}. Using chordal graphs makes the problems much smaller than directly ruling out cycles, while encoding the same restrictions. We believe that this technique is applicable in other problems using SAT-solvers as well.

>>>>

The figure shows \emph{production dependency graphs}, one for every production of the grammar. A set of semantic function definitions are given by the programmer for every production. The definitions specify how the synthesized attributes of a nonterminal are calculated in each of the different productions of that nonterminal. In semantic function definitions the inherited attributes of the nonterminal can be used as well as the synthesized attributes of its children. The attributes that are available to use, coloured gray in the picture, we call the \emph{input} attributes. Besides the semantic functions of the synthesized attributes of the parent node, the semantic functions of the inherited attributes of the children need to be defined. The attributes that require a definition, coloured white in the picture, we call the \emph{output} attributes. Note that with a semantic function definition for all output attributes we know how to compute all attributes as every input attribute is an output attribute of another node\footnote{Except for the inherited attributes of the root node. Its values need to be given as an argument to the semantic evaluator}.

Although we talk about dependency graphs in the context of scheduling, we actually draw the edges in the other direction. The edges in the picture represent data flow, which from an attribute grammar perspective is much more intuitive. For example, in the \emph{Bin} production the attribute a1 of the child \emph{r} is computed from a1 from the parent and a3 from the child \emph{l}. The edges are thus always directed from an input attribute to an output attribute, and the actual dependency graph can be obtained by reversing all edges. The edges that follow directly from the source code are the \emph{direct dependencies}.

@@

The algorithm is guaranteed to find the global optimum. For our largest example, the solution found had a total of 130 visits, which was 29 visits less in total than the previously known optimimum, found using backtracking heuristics.

@@

To compute this, we introduce a SAT variable $E(a)$ for every attribute $a$, that expresses whether or not $a$ is in $E$ or not. We set $E(a)$ to be true for the initial eager attribute $a$. We go over the graphs for the nonterminals and production rules, and generates constraints that express that whenever $a$ points to $b$ and we have $E(b)$, then we also need $E(a)$.

@@

We have run this algorithm on ever output attribute of the top-level nonterminal of all our examples. For our largest grammar, the hardest output to compute took 1 second. So, while a harder optimisation than the previous one, it is very doable in practice.

@@

In the translation into the SAT-problem we have used a novel technique for ruling out cycles in the SAT problem using chordal graphs. The technique was inspired by existing work on chordality for expressing equality logic \cite{bryant-chordal}. Using chordal graphs makes the problems much smaller than directly ruling out cycles, while encoding the same restrictions. We believe that this technique is applicable in other problems using SAT-solvers as well.
